$p$-adic precision

### 2014-02-15 padicprec.tex
Working on Section 2

\newcommand{\val}{\text{\rm val}}
\newcommand{\pr}{\text{\rm pr}}
\newcommand{\tr}{\text{\rm Tr}}
\newcommand{\com}{\text{\rm Com}}
\newcommand{\Grass}{\text{\rm Grass}}
\renewcommand{\prec}{\text{\rm prec}}

\newcommand{\lb}{\ensuremath{\llbracket}}
\newcommand{\rb}{\ensuremath{\rrbracket}}

@@

Let $K$ be a non-Archimedean local field with absolute value
$|\cdot| : K \to \R_+$ and value group $p^\Z$.   For example, we may take $K = \Qp$ with the
$p$-adic absolute value or $K = \Fp\lb t \rb$ with the $t$-adic absolute value.
Write $\OK$ for the ring of integers of $K$ and choose a dense subring $R \subset K$
consisting of elements that can be represented exactly.  For $K = \Qp$ we may
choose $R = \Z[\frac1p]$ or $R = \Q$; for $K = \Fp\lb t \rb$ we may choose
$R = \Fp[t, t^{-1}]$ or $R = \Fp(t)$.  Choose a collection of subsets $S_h \subset R$
indexed by positive integers $h$ so that
\begin{enumerate}
\item $S_h \subset S_{h+1}$,
\item $\{B_K(x, p^{-h}) : x \in S_h\}$ disjointly covers $K$.
\end{enumerate}
The choice of the $S_h$ is possible since $K$ is non-Archimedean and locally compact.

An \emph{absolute value} on a field $K$ is a multiplicative map $|\cdot| 
: K \to \R_+$ satisfying the triangle inequality and sending every 
nonzero element of $K$ to a strictly positive real number. It is said 
\emph{ultrametric} if it satisfies the following strong from of the 
triangle inequality:
$$|x+y| \leq \max(|x|, |y|).$$
Given an absolute value $|\cdot|$ on $K$, we note that $|K^\times|$ is a 
multiplicative subgroup of $\R_+^\times$. It is then either discrete or 
dense. We shall say that $|\cdot|$ is \emph{discrete} if $|K^\times|$ is
discrete. The valued field $(K, |\cdot|)$ is said \emph{complete} if all 
Cauchy sequences (with respect to the distance $d$ defined by $d(x,y) = 
|x-y|$) converge.

The formula $|x| = p^{-\val(x)}$ defines an ultrametric absolute value on 
the field $\Qp$ for which it is complete. More generally, the $p$-adic 
absolute value extends uniquely to any finite extension $K$ of $\Qp$ and 
turns $K$ into a complete ultrametric field. Another classical example of 
complete ultrametric field is the field of Laurent series over a finite
field, the absolute value being defined from the valuation of a series.

From now, we fix a field $K$ equipped with an ultrametric absolute value 
denoted by $|\cdot|$ and we assume that $K$ is complete. Given $v$ a 
$K$-vector space (of possibly infinite dimension), an \emph{ultrametric 
norm} on $V$ is a map $\Vert\cdot\Vert : V \to \R^+$ satisfying:
\begin{enumerate}[(i)]
\item $\Vert x\Vert = 0$ if and only if $x = 0$;
\item $\Vert \lambda x\Vert = |\lambda| \cdot \Vert x\Vert$;
\item $\Vert x+y\Vert \leq \max(\Vert x\Vert, \Vert y\Vert)$.
\end{enumerate}
A norm gives rise to a distance and a normed $K$-vector space $V$ is 
said \emph{complete} when any Cauchy sequence with values in $V$ 
converges. A normed $K$-vector space which is complete is called a 
\emph{$K$-Banach}.

Let $E$ be a $K$-normed vector space. Given $r \in \R^+$, 
we denote by $B^-_E(r)$ (resp. by $B_E(r)$) the set of elements $x 
\in E$ such that $\Vert x \Vert < r$ (resp. $\Vert x \Vert \leq r$). 
We insist on the following maybe unusual fact: $B^-_E(r)$ and 
$B_E(r)$ are both open and closed for the norm topology on $E$.
Clearly, $B_K(0,1)$ is stable under multiplication. Moreover, the 
ultrametric triangle inequality ensures that $B_K(1)$ is stable under 
addition as well. Thus, $B_K(1)$ is a subring of $K$: it is often 
called the \emph{integer ring} of $K$.

\begin{deftn}
Let $E$ be a $K$-Banach.
A \emph{lattice} in $E$ is an open bounded sub-$B_K(1)$-module of $E$.
\end{deftn}

\begin{rmk}
Any lattice $H$ in $E$ is also closed since its complement is the
union of all cosets $a + H$ (with $a \not\in H$) which are all open.
\end{rmk}

@@

Let $f : \Qp^2 \to \Qp^2$ be the function mapping $(x,y)$ to $(x+y, 
x-y)$. We consider in addition two $p$-adic numbers $x$ and $y$, 
which are known up to precision $O(p^n)$ and $O(p^m)$ respectively.
The question we would like to address in this example is the
computation of $f \circ f(x,y)$.

@@

The theory of $p$-adic precision we want to develop is mainly based on a 
single lemma (Lemma \ref{lem:main}) of $p$-adic analysis. The aim of 
this Section is to state and prove this lemma. Consequences will be 
discussed after in Section \ref{sec:tracking}.

@@

%The following Lemma shows that the notion of lattices behaves rather 
%well under linear\footnote{We shall see in the sequel (\emph{cf} Lemma 
%\ref{lem:main}) that, under some extra hypothesis, the linearity 
%assumption can be drastically relaxed.} transformations.
%
%\begin{lem}
%\label{lem:morlat}
%Let $E$ and $F$ be two $K$-Banach and $f : 
%E \to F$ be a continuous \emph{surjective} linear map.
%If $H$ is a lattice in $E$, then $f(H)$ is a lattice in $F$.
%\end{lem}
%
%\begin{proof}
%It is clear that $f(H)$ is a module over $B_K(1)$ which is bounded.
%The fact that it is open is a direct consequence of the open mapping
%Theorem.
%\end{proof}
%
%To each lattice $H$ in $E$, one can attach a norm $\Vert \cdot \Vert_H$ 
%on $E$ defined as follows. We pick $x \in E$, $x \neq 0$. We introduce
%the set $I_x$ consisting of scalars $\lambda \in K$ such that $\lambda x 
%\in H$. It follows from the fact that $H$ is open and bounded in $E$
%that $I_x$ is nonempty and bounded in $K$. Therefore, the image $|I_x|$ 
%of $I_x$ under the absolute value map is bounded and contains a positive
%element. We can then define:
%$$\Vert x \Vert_H = (\sup |I_x|)^{-1}.$$
%It is an exercise to check that $\Vert \cdot \Vert_H$ is a norm over 
%$E$.
%
%\begin{lem} \label{lem:reductionball}
%\begin{enumerate}
%\item The norms $\Vert \cdot \Vert_H$ and $\Vert \cdot \Vert$ are 
%equivalent.
%\item Let $B^-_{E,H}(1)$ (resp. by $B_{E,H}(1)$) the set of elements $x 
%\in E$ such that $\Vert x \Vert_H < 1$ (resp. $\Vert x \Vert_H \leq 1$).
%Then:
%$$B^-_{E,H}(1) \subset H \subset B_{E,H}(1)$$
%and the last inclusion is an equality if the absolute value on $K$ is
%discrete.
%\end{enumerate}
%\end{lem}
%
%\begin{proof}
%Left to the reader.
%\end{proof}
%
%The first part of Lemma \ref{lem:reductionball} asserts that the 
%identity map $\iota_H : (E, \Vert \cdot \Vert_H) \to (E, \Vert \cdot 
%\Vert)$ is an homeomorphism. Furthermore, by the second of the Lemma, 
%$H$ is the image under $\iota_H$ of a lattice satisfying an extra 
%assumption: it is ``between the open unit ball and the closed unit 
%ball''.

@@

\begin{deftn}
\label{def:firstorder}
Let $E$ and $F$ be two $K$-Banach and $f : U \rightarrow F$ be a 
function defined on an open subset $U$ of $E$. Let also $v_0$ be a 
point in $U$.
A lattice $H$ in $E$ is called a \emph{first order lattice} for $f$ at 
$v_0$ if the equality
$$f(v_0 + H) = f(v_0) + f'(v_0) (H)$$
holds.
\end{deftn}

@@

Then, for all $\rho \in (0, 1]$, there exists a positive real 
number $\delta$ such that, for all $r \in (0, \delta)$, any lattice
$H$ such that $B^-_E(\rho r) \subset H \subset B_E(r)$ is a first
order lattice for $f$ at $v_0$.
\end{lem}

\begin{proof}
We may assume that $v_0=0$ and $f(0)=0$. Since $f'(0)$ is surjective, by 
the open mapping Theorem, there exists $C>0$ such that $B_F(1) \subset 
C  f'(0)(B_E(1))$. Let $\varepsilon>0$ be such that 
$\varepsilon C < \rho$. Let $U_\varepsilon \subset E$ given by Definition 
\ref{deftn : diff} and the fact that $f$ is differentiable at $0$. We 
may assume $U_\varepsilon = B_E(\delta)$ for some $\delta >0$.

Let $r \in (0, \delta)$. Let $B^-_E(r) \subset H \subset B_E(r).$ Let us 
show that $f$ maps surjectively $H$ onto $f'(0) (H)$. We first prove 
that $f(H) \subset f'(0) (H)$. Let $x \in H$. By differentiability at 
$0$, $\Vert f(x)-f'(0)(x) \Vert \leq \varepsilon \Vert x \Vert $. 
Therefore, if $y=f(x)-f'(0)(x)$, then $\Vert y \Vert \leq 
\varepsilon r$. By definition of $C$, we have $B_F(\varepsilon r) 
\subset \varepsilon C f'(0) (B_E(r))$. Thus there exists $x_1 \in 
\varepsilon C \cdot B_E(r)$ such that $f'(0) (x_1) =y$. Since 
$\varepsilon C < \rho$, we get $x_1 \in B^-_E(\rho r) \subset H$ and 
then $f(x)= f'(0) (x-x_1) \in f'(0) (H)$.

We shall now prove surjectivity. Let $y \in f'(0) (H)$. Let $x_0 \in H$ 
be such that $y = f'(0) (x_0)$. We define inductively two sequences 
$(x_n)$ and $(z_n)$ as follows:
\begin{itemize}
\item $z_n$ is an element of $E$ satisfying $f'(0)(z_n) = y - 
f(x_n)$ and $\Vert z_n \Vert \leq C \cdot \Vert y - f(x_n) \Vert$ (such 
an element exists by definition of $C$), and
\item $x_{n+1}=x_n+z_n$.
\end{itemize}
For convenience, let us also define $x_{-1} = 0$. We claim that the 
sequences $(x_n)$ and $(z_n)$ are well defined and take their values in 
$H$. To establish this, it is enough to prove that, assuming that 
$x_{n-1}$ and $x_n$ belongs to $H$, so do $z_n$ and $x_{n+1}$. Noticing
that 
\begin{equation}
\label{eq:mainlemma}
y - f(x_n) 
= f(x_{n-1}) + f'(0)(z_{n-1}) - f(x_n)
= f(x_{n-1}) - f(x_n) - f'(0)(x_{n-1} - x_n)
\end{equation}
we deduce using differentiability that
$\Vert y - f(x_n) \Vert \leq \varepsilon \cdot \Vert x_n - x_{n-1}
\Vert$.
Since we are assuming that $x_{n-1}$ and $x_n$ lie in $H \subset
B_E(r)$, we find $\Vert y - f(x_n) \Vert \leq \varepsilon r$. Thus
$\Vert z_n \Vert \leq C \cdot \varepsilon r < \rho r$ and then
$z_n \in H$. From the relation $x_{n+1} = x_n + z_n$, we finally
deduce $x_{n+1} \in H$.

Now using again Eq.~\eqref{eq:mainlemma} and differentiability at $0$,
we get, for all positive integer $n$,
$$\Vert y  - f(x_n) \Vert 
\leq \varepsilon \cdot \Vert z_{n-1} \Vert \leq \varepsilon C \cdot \Vert 
y - f(x_{n-1}) \Vert,$$
from what we deduce that $\Vert y - f(x_n) \Vert = O(a^n)$ and 
$\Vert z_n \Vert = O(a^n)$ with $a = \varepsilon C < \rho \leq 1$. 
This shows that $(x_n)$ is a Cauchy sequence and, since $E$ is complete, 
it converges. Because $H$ is closed, if $x$ denotes the limit of 
$(x_n)$, we have $x \in H$. Moreover, as a consequence of the 
differentiability, $f$ is continuous over $U_\varepsilon$. Therefore, 
since $x \in H \subset U_\varepsilon$, we get $y=f(x)$ and we are done.
\end{proof}

Let us conclude this paragraph by a remark on the assumption made in 
Lemma \ref{lem:main} about the surjectivity of $f'(v_0)$. First let us 
emphasize that this hypothesis is definitely necessary. Indeed, it would 
implies otherwise that the image of $f$ is locally contained in a proper 
sub vector space around each point where the differential of $f$ is not 
surjective, which is certainly not true! Nevertheless, one can get 
around this issue in many situations. Indeed, consider the situation of 
Lemma \ref{lem:main} without assuming that $f'(v_0)$ is surjective. Now, 
we pick a closed sub vector space $K$ of $F$ such that $K + f'(v_0)(E) = 
F$. Then, denoting by $\pr_K$ the canonical projection of $F$ onto 
$F/K$, the composite $\pr_K \circ f$ is certainly differentiable at 
$v_0$ at its differential is surjective. Hence Lemma \ref{lem:main}
applies and shows that
\begin{equation}
\label{eq:notsurjective}
f(v_0 + H) \subset f(v_0) + f'(v_0)(H) + K
\end{equation}
for all lattices $H$ satisfying a certain condition. Now we can take
the intersection of the right hand side of \eqref{eq:notsurjective} 
over all possible $K$ and hope that we get this way a sharp affine upper 
bound of $f(v_0 + H)$.

@@

%Using Lemma \ref{lem:reductionball}, one can extend Proposition 
%\ref{prop:locanalytic} to arbitrary lattices $H$ as follows. Given $H$
%a lattice in some $K$-Banach space $E$, we first define the two
%following quantities:
%\begin{itemize}
%\item $\max(H)$ is the smallest $r$ such that $H \subset B_E(r)$, and
%\item $\min(H)$ is the greatest $r$ such that $B_E(r) \subset H$.
%\end{itemize}
%
%\begin{cor}
%Let $U$ are $V$ be open subsets in $K$-Banach spaces $E$ and $F$ 
%respectively and $f : U \to V$ be a locally analytic function. Let 
%also $v_0$ be a point in $U$ such that $f'(v_0)$ is surjective. 
%Then a lattice $H$ in $E$ is a first order lattice for $f$ at $v_0$ as 
%soon as $\max(H)^2 < C_{f, v_0} \cdot \min(H)$ where $C_{f, v_0}$ is a 
%constant depending only on $f$ and $v_0$ which can be determined 
%explicitely from the Newton polygon of $f$ at $v_0$.
%\end{cor}
%
%\begin{proof}
%We apply Proposition \ref{prop:locanalytic} to the composite $f \circ 
%\iota_H$ where $\iota_H : (E, \Vert \cdot \Vert_H) \to (E, \Vert \cdot 
%\Vert)$ is induced by identity map.
%\end{proof}

@@

With above definitions, Lemma \ref{lem:main} extends readily to 
manifolds. To state it the context of manifolds, we first need to define 
a norm on the tangent space $T_x V$ (where $V$ is some $K$-variety and 
$x$ is a point in $V$). There is actually in general no canonical choice 
for this. Indeed, let us consider a $K$-manifold $V$ covered by charts 
$U_i$'s ($i \in I$) which are open subset of $K$-Banach $E_i$'s. If $x$ 
is a point in $V$, the tangent space $T_x V$ is by definition isomorphic 
to $E_i$ for each index $i$ such that $x \in V_i$. A natural norm on 
$T_x V$ is then the one obtained by pulling back the norm on $E_i$. 
However, since the transition maps are not required to be isometries, 
this norm do depend on the choice on $i$. They are nevertheless all
equivalent because the transition maps are required to be continuous.

In the next Lemma, we choose to endow $T_x V$ by one of the above norms.

\begin{lem}
Let $V$ and $W$ be two $K$-manifolds. 
Suppose that we are given a differentiable function $f : V \to W$, 
together with a point $x \in V$ such that $f'(x) : T_x V \to T_{f(x)} W$ 
is surjective.

Then for all $\rho \in (0,1]$, there exists $\delta \in (0,1)$ such 
that for all $r \in (0, \delta)$, the equality
$$f(x + O(H)) = f(x) + O(f'(x)(H))$$
holds for all lattices $H$ in $T_x V$ such that $B^{\phantom -}_{T_x V}(\rho r)
\subset H \subset B^-_{T_x V}(r)$.
\end{lem}

@@

\begin{rmk}
The constant $\delta$ that appears in the above Lemma depends (up to
some multiplicative constant) on the norm that we have chosen on $T_x V$. 
However, once this norm is fixed, and assuming further that $V$ and $W$ 
are locally analytic $K$-manifolds and the mapping $f$ is locally 
analytic as well, the constant $\delta$ can be made explicit using the 
method of \S \ref{ssec:locanalytic}.
\end{rmk}

>>>>

\newcommand{\val}{\text{\rm val}}
\newcommand{\tr}{\text{\rm Tr}}
\newcommand{\com}{\text{\rm Com}}
\newcommand{\Grass}{\text{\rm Grass}}
\renewcommand{\prec}{\text{\rm prec}}

\newcommand{\lb}{\ensuremath{\llbracket}}
\newcommand{\rb}{\ensuremath{\rrbracket}}
\newcommand{\lp}{(\!(}
\newcommand{\rp}{)\!)}

@@

Let $K$ be a field with absolute value $|\cdot| : K \to \R_{\ge 0}$. 
We assume that the induced metric is an ultrametric -- 
\[ \lvert x + y \rvert \leq \max(\lvert x \rvert, \lvert y \rvert) \]
-- and that $K$ is complete with respect to it.
For example, we may take $K = \Qp$ with the
$p$-adic absolute value or $K = k\lp t \rp$ with the $t$-adic absolute value.
Write $\OK$ for the ring $\{x \in K : \lvert x \rvert \le 1\}$ and assume
that $K$ contains a dense subring $R \subset K$
consisting of elements that can be represented with a finite amount of space.
For $K = \Qp$ we may choose $R = \Z[\frac1p]$ or $R = \Q$;
for $K = \Fp\lb t \rb$ we may choose $R = \Fp[t, t^{-1}]$ or $R = \Fp(t)$.

If $E$ is a $K$-vector space, possibly of infinite dimension, then an
\emph{ultrametric norm} on $E$ is a map $\Vert\cdot\Vert : E \to \R^+$ satisfying:
\begin{enumerate}[(i)]
\item $\Vert x\Vert = 0$ if and only if $x = 0$;
\item $\Vert \lambda x\Vert = |\lambda| \cdot \Vert x\Vert$;
\item $\Vert x+y\Vert \leq \max(\Vert x\Vert, \Vert y\Vert)$.
\end{enumerate}
A \emph{$K$-Banach space} is a complete normed $K$-vector space.

\begin{deftn} % Separated definitions?
Let $E$ be a $K$-Banach space.
A \emph{lattice} in $E$ is an open bounded sub-$\OK$-module of $E$.
\end{deftn}

For a $K$-Banach space $E$ and $r \in \R_{\ge 0}$, write
\begin{align*}
B_E(r) &= \{ x \in E : \Vert x \Vert \le r \}, \\
B^-_E(r) &= \{ x \in E : \Vert x \Vert < r\}.
\end{align*}
Note that $B_E(r)$ and $B^-_E(r)$ are both lattices.

\begin{deftn}
Suppose $E$ is a $K$-Banach space and $I$ a set.
A family $(x_i)_{i \in I} \subset E$ is a \emph{Banach basis} for $E$ if
every element $x \in E$ can be written
\[
x = \sum_{i \in I} \alpha_i x_i
\]
for scalars $\alpha_i \in K$ with $\alpha_i \to 0$, and $\Vert x \Vert = \sup_{i \in I} \lvert \alpha_i \rvert$.
\end{deftn}
Note that if $E$ is finite dimensional then the condition $\alpha_i \to 0$ is vacuous.
% Lattice from Banach basis

\subsection{Approximate elements}

Suppose that $E$ is a $K$-Banach space with basis $(x_i)_{i \in I}$.  We say that an element
$x \in E$ is \emph{exact} if there is a finite subset $J \subseteq I$ and scalars $\alpha_j \in R$ with
\begin{equation} \label{eq:exact_elt}
x = \sum_{j \in J} \alpha_j x_j.
\end{equation}
The set of exact elements is dense in $E$.

An approximate element is a pair $(x, H)$ where $x \in E$ is an exact element
and $H$ is a lattice in $E$.  Such a pair represents an undetermined element of the coset
$x + H$.  We will frequently write $x + O(H)$ to emphasize the fact that $H$ represents
the uncertainty in the value of the approximate element.  In the special case that $E = K = \Qp$,
we recover the standard notation $a + O(p^n)$ for an approximate $p$-adic element.

Note that using an arbitrary lattice to represent the precision of an approximate element
can reduce precision loss when compared to storing the precision of each coefficient
$\alpha_i$ in \eqref{eq:exact_elt} separately.  Consider the function 
$f : \Qp^2 \to \Qp^2$ mapping $(x,y)$ to $(x+y, x-y)$ and the problem of computing
$f \circ f(a + O(p^n), b + O(p^m))$.

@@

Let

@@

The theory of $p$-adic precision we want to develop is mainly based 
on a single lemma of $p$-adic analysis. The aim of this Section is
to state and prove this lemma. Consequences will be discussed after
in Section \ref{sec:tracking}.

@@

\subsection{Ultrametric analysis}
\label{ssec:ultrametric}

The following Lemma shows that the notion of lattices behaves rather 
well under linear\footnote{We shall see in the sequel (\emph{cf} Lemma 
\ref{lem:main}) that, under some extra hypothesis, the linearity 
assumption can be drastically relaxed.} transformations.

\begin{lem}
\label{lem:morlat}
Let $E$ and $F$ be two $K$-Banach and $f : 
E \to F$ be a continuous \emph{surjective} linear map.
If $H$ is a lattice in $E$, then $f(H)$ is a lattice in $F$.
\end{lem}

\begin{proof}
It is clear that $f(H)$ is a module over $B_K(1)$ which is bounded.
The fact that it is open is a direct consequence of the open mapping
Theorem.
\end{proof}

To each lattice $H$ in $E$, one can attach a norm $\Vert \cdot \Vert_H$ 
on $E$ defined as follows. We pick $x \in E$, $x \neq 0$. We introduce
the set $I_x$ consisting of scalars $\lambda \in K$ such that $\lambda x 
\in H$. It follows from the fact that $H$ is open and bounded in $E$
that $I_x$ is nonempty and bounded in $K$. Therefore, the image $|I_x|$ 
of $I_x$ under the absolute value map is bounded and contains a positive
element. We can then define:
$$\Vert x \Vert_H = (\sup |I_x|)^{-1}.$$
It is an exercise to check that $\Vert \cdot \Vert_H$ is a norm over 
$E$.

\begin{lem} \label{lem:reductionball}
\begin{enumerate}
\item The norms $\Vert \cdot \Vert_H$ and $\Vert \cdot \Vert$ are 
equivalent.
\item Let $B^-_{E,H}(1)$ (resp. by $B_{E,H}(1)$) the set of elements $x 
\in E$ such that $\Vert x \Vert_H < 1$ (resp. $\Vert x \Vert_H \leq 1$).
Then:
$$B^-_{E,H}(1) \subset H \subset B_{E,H}(1)$$
and the last inclusion is an equality if the absolute value on $K$ is
discrete.
\end{enumerate}
\end{lem}

\begin{proof}
Left to the reader.
\end{proof}

The first part of Lemma \ref{lem:reductionball} asserts that the 
identity map $\iota_H : (E, \Vert \cdot \Vert_H) \to (E, \Vert \cdot 
\Vert)$ is an homeomorphism. Furthermore, by the second of the Lemma, 
$H$ is the image under $\iota_H$ of a lattice satisfying an extra 
assumption: it is ``between the open unit ball and the closed unit 
ball''.

@@

@@

Then, there exists a positive real number $\delta$ such that, for all $r 
\in (0, \delta)$, the equality
\begin{equation}
\label{eq:main}
f(v_0 + H) = f(v_0) + f'(v_0) (H).
\end{equation}
holds for each lattice $H$ such that $B^-_E(r) \subset H \subset
B_E(r)$.
\end{lem}

\begin{rmk}
\label{rmk:mainlemma}
Given a lattice $H_0$ in $E$, the equality \eqref{eq:main} holds for $H 
= \lambda H_0$ with $\lambda$ small enough. To prove this, it is enough 
to apply Lemma \ref{lem:main} to the map $f \circ \iota_{H_0}$ where 
$\iota_{H_0} : (E, \Vert \cdot \Vert_{H_0}) \to (E, \Vert \cdot \Vert)$ 
is the ``identity map'' (which is an homeomorphism by Lemma 
\ref{lem:reductionball}).
\end{rmk}

\begin{proof}
We may assume that $v_0=0$ and $f(0)=0$. Since $f'(0)$ is surjective, by 
the open mapping Theorem, there exists $C>0$ such that $B_F(1) \subset 
C  f'(0) \cdot B_E(1)$. Let $\varepsilon>0$ be such that 
$\varepsilon C < 1$. Let $U_\varepsilon \subset E$ be given by by definition \ref{deftn : diff} and the fact that $f$ is differentiable at $0$. We may assume $U_\varepsilon = 
B_E(\delta)$ for some $\delta >0$.

Let $r \in (0, \delta)$. Let $B^-_E(r) \subset H \subset
B_E(r).$
Let us show that $f$ maps surjectively $H$ onto $f'(0) (H)$.

We first prove that $f(H) \subset f'(0) (H)$.
Let $x \in H$. By differentiability at $0$, $\Vert f(x)-f'(0) \cdot x \Vert \leq 
\varepsilon \Vert x \Vert $. Therefore, if $y=f(x)-f'(0) \cdot x,$ then $\Vert y \Vert \leq \varepsilon r$. As a consequence of the open mapping Theorem, $B_F(\varepsilon r) \subset 
\varepsilon C  f'(0) \cdot B_E(r),$ and there exists $x_1 \in \varepsilon C  \cdot B_E(r)$ such that $f'(0) \cdot x_1 =y.$ Since $\varepsilon C <1,$  $x_1 \in B^-_E(r) \subset H.$ and finally, $f(x)= f'(0) \cdot (x-x_1) \in f'(0) (H)$.

We shall now prove surjectivity. Let $y \in f'(0) (H)$.
Let $x_0 \in H$ be such that $y = f'(0) \cdot x_0$. Since $f(x_0) \in f'(0) (H)$, we can define $z_0 \in E$ to be such that $f'(0) \cdot z_0 = y - f(x_0)$ and $\Vert z_0 \Vert \leq C \cdot \Vert y - f(x_0) \Vert$.
We then define by induction two sequences $(x_n)$ and $(z_n)$ by
\begin{itemize}
\item $x_{n+1}=x_n+z_n$ and 
\item $z_n \in E$ is such that
$f'(0) \cdot z_n = y - f(x_n)$ and
$\Vert z_n \Vert \leq C \cdot \Vert y - f(x_n) \Vert$.
\end{itemize}
By definition, $\Vert z_0 \Vert < C \Vert y_0\Vert$ and $y 
= -f(x_0)+f'(0) \cdot x_0$. Therefore, $\Vert y \Vert < \varepsilon 
\Vert x_0 \Vert \leq \varepsilon r$. Thus $\Vert z_0 \Vert \leq 
\varepsilon C \cdot r < r$. Thus, $z_0$ and $x_1=x_0+z_0$  lie in $B^-_E(r) 
\subset H$. An easy induction involving an analoguous computation
shows that the sequences $(x_n)$ and $(z_n)$ are well defined and
take their values in $H$.

We notice that $y - f(x_{n+1}) = f(x_n) - f(x_{n+1}) + f'(0) \cdot z_n$.
Since $z_n=x_{n+1}-x_n,$, if we use differentiability, we get:
$$\Vert y  - f(x_{n+1}) \Vert 
\leq \varepsilon \Vert z_n \Vert \leq \varepsilon C \cdot \Vert 
y - f(x_n) \Vert,$$
from what we deduce that $\Vert y - f(x_{n+1}) \Vert = 
O(a^n)$ and $\Vert z_n \Vert = O(a^n)$ with $a = \varepsilon C < 1$.
This shows that $(x_n)$ is a Cauchy sequence and, since $E$ is complete, it converges. Because $H$ is closed, if $x$ denotes the limit of $(x_n)$, we have $x \in H.$ Yet, as a consequence of the differentiability, $f$ is continuous over $U_\varepsilon.$ Therefore, since $x \in H \subset U_\varepsilon,$ then  $y=f(x)$. The result is proven.
\end{proof}

\begin{deftn}
\label{def:firstorder}
Let $E$ and $F$ be two $K$-Banach and $f : U \rightarrow F$ be a 
function defined on an open subset $U$ of $E$. Let also $v_0$ be a 
point in $U$.
A lattice $H$ in $E$ is called a \emph{first order lattice} for $f$ at 
$v_0$ if
$$f(v_0 + H) = f(v_0) + f'(v_0) (H).$$
\end{deftn}

By Lemma \ref{lem:main} and Remark \ref{rmk:mainlemma}, if $f$ is 
differentiable at $v_0$ and $f'(v_0)$ is surjective, then for all 
lattice $H$ in $E$, there exists a constant $\delta$ such that $aH$ 
is a first order lattice for $f$ at $v_0$ for all $a \in B_E(r)$.

@@

Using Lemma \ref{lem:reductionball}, one can extend Proposition 
\ref{prop:locanalytic} to arbitrary lattices $H$ as follows. Given $H$
a lattice in some $K$-Banach space $E$, we first define the two
following quantities:
\begin{itemize}
\item $\max(H)$ is the smallest $r$ such that $H \subset B_E(r)$, and
\item $\min(H)$ is the greatest $r$ such that $B_E(r) \subset H$.
\end{itemize}

\begin{cor}
Let $U$ are $V$ be open subsets in $K$-Banach spaces $E$ and $F$ 
respectively and $f : U \to V$ be a locally analytic function. Let 
also $v_0$ be a point in $U$ such that $f'(v_0)$ is surjective. 
Then a lattice $H$ in $E$ is a first order lattice for $f$ at $v_0$ as 
soon as $\max(H)^2 < C_{f, v_0} \cdot \min(H)$ where $C_{f, v_0}$ is a 
constant depending only on $f$ and $v_0$ which can be determined 
explicitely from the Newton polygon of $f$ at $v_0$.
\end{cor}

\begin{proof}
We apply Proposition \ref{prop:locanalytic} to the composite $f \circ 
\iota_H$ where $\iota_H : (E, \Vert \cdot \Vert_H) \to (E, \Vert \cdot 
\Vert)$ is induced by identity map.
\end{proof}

@@

With above definitions, Lemma \ref{lem:main} extends readily to 
manifolds.

\begin{lem}
Let $V$ and $W$ be two $K$-manifolds. 
Suppose that we are given a differentiable function $f : V \to W$, 
together with a point $x \in V$ such that $f'(x) : T_x V \to T_{f(x)} W$ 
is surjective. Let $H$ be a precision datum at $x$. Then, there exists
a positive real number $\delta$ such that, for all $\lambda \in 
B_K(\delta)$, the equality
$$f(x + \lambda H) = f(x) + \lambda \cdot f'(x)(H)$$
holds.
\end{lem}

@@

\begin{rmk}
If $V$ and $W$ are locally analytic $K$-manifolds and the mapping $f$ is 
locally analytic as well, then one can explicitely determine, using the 
method of \S \ref{ssec:locanalytic}, a value $\delta$ for which the 
above Lemma works.
\end{rmk}
